Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-03-27 paper/high-assurance-contracts.tex
replace wellformed by Pascalish. Doesn't sound quite right but will do for now

How does this notion of correctness integrate with the verification
conditions generated by AGREE? In \emph{contract correctness}, the
model checker is proving the following property
\[
\konst{Hist}(A) \imp P
\]
and in \emph{code contract correctness} we essentially prove a Hoare
triple of the form
\[
\set{\konst{Hist}(A)}\; (\mathit{Eqs} \cdot G) \; \set{P}
\]
This accords with intuition, namely that AGREE is a kind of
`program-free program logic'; by pulling out a program $(\mathit{Eqs}
\cdot G)$ from a code contract we find ourselves in the domain of imperative program verification. For example, at this
point one could apply a verification condition generator to help break
down the proof obligation. In order to use the code contract result in
system contract verification conditions, one has to show that the code
contract result implies the system verification condition.

\subsubsection*{Wellformedness}
Stream-based evaluation allows values arbitrarily `deep in the past'
to be accessed by use of $\konst{pre}(-)$. However, the programming
languages being targeted by SPLAT do not directly support streams.  We
are currently developing a means of translating away such deep
temporal accesses by the introduction of intermediate variables. For
now, we focus on the notion of \emph{wellformed} equations that needs
to hold. The guiding intuition is that a wellformed {\Eqs} allows
access at most one step in the past, and accesses of a past value for
a variable are only possible until the variable is updated. In
informal terms, this means that ${\Eqs}\cdot G$ in a code contract
must be such that:
\begin{itemize}
\item Followed-by only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
\item $\konst{pre}(v)$ is allowed only before the defining equation $v = e$ (and may occur in $e$).
\end{itemize}
For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
defined, albeit in a non-wellformed manner, by
{\small
\begin{lstlisting}[style=agree]
  Fib = 1 -> pre(1 -> Fib + pre Fib)
\end{lstlisting}
}
\noindent or by (again non-wellformed)
{\small
\begin{lstlisting}[style=agree]
  N = 0 -> 1 + pre(N)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib))
\end{lstlisting}
}
\noindent or by (wellformed):
{\small
\begin{lstlisting}[style=agree]
  N = 0 -> 1 + pre(N)
  F2 = 42 -> pre(F1)
  F1 = 42 -> pre(Fib)
  Fib = if N <= 1 then 1 else F1 + F2
\end{lstlisting}
}
A wellformed ${\Eqs} \cdot G$ is straightforward to implement in
conventional programming languages, as we will see in
Section \ref{sec:synthesis}.

>>>>

How does this notion of correctness integrate with the verification
conditions generated by AGREE? In \emph{contract correctness}, the
model checker is proving the following property
\[
\konst{Hist}(A) \imp P
\]
and in \emph{code contract correctness} we essentially prove a Hoare
triple of the form
\[
\set{\konst{Hist}(A)}\; (\mathit{Eqs} \cdot G) \; \set{P}
\]
This accords with intuition, namely that AGREE is a kind of
`program-free program logic'; by pulling out a program
$(\mathit{Eqs}\cdot G)$ from a code contract we find ourselves in the
domain of imperative program verification. For example, at this
point one could apply a verification condition generator to help break
down the proof obligation. In order to use the code contract result in
system contract verification conditions, one has to show that the code
contract result implies the system verification condition.

\subsubsection*{Imperative code contracts}
Stream-based evaluation allows values arbitrarily `deep in the past'
to be accessed by use of $\konst{pre}(-)$. However, the programming
languages targeted by SPLAT do not directly support streams.  We are
developing a translation to eliminate such deep temporal accesses by
the introduction of intermediate variables. The essential property the
translation needs to satisfy is a syntactic restriction ---call
it \emph{Pascalish}---that supports easy translation to standard
imperative languages.  The guiding intuition is that a Pascalish
{\Eqs} allows access at most one step in the past, and accesses to a
past value of a variable are only possible until the variable is
updated. In informal terms, this means that ${\Eqs}\cdot G$ in a code
contract has the following properties:
\begin{itemize}
\item Followed-by only occurs at the top level of a variable definition, \eg, $v = e_1 \to e2$;
\item $\konst{pre}(e)$ is allowed only when $e$ is a variable; and
\item $\konst{pre}(v)$ is allowed only before the defining equation $v = e$ (and may occur in $e$).
\end{itemize}
For example, the Fibonacci sequence $1,1,2,3,5,8,13,\ldots$ can be
defined, albeit in a non-Pascalish manner, by
{\small
\begin{lstlisting}[style=agree]
  Fib = 1 -> pre(1 -> Fib + pre Fib)
\end{lstlisting}
}
\noindent or by (again non-Pascalish)
{\small
\begin{lstlisting}[style=agree]
  N = 0 -> 1 + pre(N)
  Fib = if N <= 1 then 1 else pre(Fib) + pre(pre(Fib))
\end{lstlisting}
}
\noindent or by (Pascalish):
{\small
\begin{lstlisting}[style=agree]
  N = 0 -> 1 + pre(N)
  F2 = 42 -> pre(F1)
  F1 = 42 -> pre(Fib)
  Fib = if N <= 1 then 1 else F1 + F2
\end{lstlisting}
} A Pascalish ${\Eqs} \cdot G$ is straightforward to implement in
conventional programming languages, as we will see in
Section \ref{sec:synthesis}.
